

\section{Fundamental Quantum Algorithms}

This section will simply be a catalogue of quantum algorithms and some of their properties. Many of these subsections will be based on the presentations given in \cite{childs17} and \cite{batty2004quantum}.

\subsection{Deutsch-Jozsa Algorithm}
Let $f: \{0,1\} \rightarrow \{0,1\}$ be a boolean function with one-bit input. We would like to ascertain if $f$ is either constant (both inputs have the same value) or balanced (inputs have different values). Through the classical perspective, one would have to query the function twice to determine the state of $f$ as constant or balanced. However, we can leverage quantum mechanics so $f$ will only have to be queried once. This is a first step towards understanding the quantum speedup through \emph{query complexity}.
%
The following circuit accomplishes this:

\begin{figure}[ht]
    \centering
    \leavevmode
    \Qcircuit {
       \ket{0} & \gate{H} & \multigate{1}{U_f} & \gate{H} & \meter \\
       \ket{1} & \gate{H} & \ghost{U_f} & \qw & \qw
    }
    \caption{Deutsch-Jozsa circuit}
\end{figure}
%
where $U_f$ is the map with the action: $\ket{x}\otimes\ket{y} \mapsto \ket{x}\otimes\ket{y \oplus f(x)}$. We calculate the action of the first two steps of the circuit above as follows:
%
\begin{gather}
  \ket{0}\ket{1} \rightarrow \frac{1}{2}(\ket{0} + \ket{1})(\ket{0} - \ket{1}) = \frac{1}{2}(\ket{0}(\ket{f(0)} - \ket{1 \oplus f(0)}) + \ket{1}(\ket{f(1)} - \ket{1 \oplus f(1)})) = \\
  \frac{1}{2}\sum_{x \in \{0,1\}} \ket{x}(\ket{f(x)} - \ket{1 \oplus f(x)}) = \\
  \frac{1}{\sqrt{2}} \sum_{x \in \{0,1\}} (-1)^{f(x)}\ket{x} \frac{1}{\sqrt{2}}(\ket{0} -
  \ket{1})
\end{gather}
By ignoring the second register, we apply the Hadamard gate to the first register to get $\ket{0}$ up to a global phase if $f$ is constant and $\ket{1}$ if $f$ is balanced. Thus, if we measure the first register, we get the respective results with certainty.

In this algorithm, we only query the $f$-oracle once instead of twice as dictated by classical intuition. In fact, we can generalize this to a $n$-bit boolean function $f:\{0,1\}^n \rightarrow \{0,1\}$ where $f$ is guaranteed to be either constant (all $n$-bit input strings map to the same value) or balanced (there are an equal number of bit strings mapping to both $0,1$).

We apply $H^{\otimes (n+1)}$ to the first register of $n$-qubits and the second result register and another iteration of $H^{\otimes n}$ after applying the $U_f$ gate accepting $n$-qubits:

\begin{gather}
\ket{0...0}\ket{1} \xrightarrow{H^{\otimes (n+1)}} \frac{1}{\sqrt{2^n}} \sum_{x = 0}^{2^n -1} \ket{x}\frac{1}{\sqrt{2}}(\ket{0}-\ket{1}) \xrightarrow{U_f} \frac{1}{\sqrt{2^n}} \sum_{x = 0}^{2^n -1} (-1)^{f(x)}\ket{x}\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})
\end{gather}

If we now discard the second result register and apply $H^{\otimes n}$ again to the first register, we yield:
$$\sum_{x = 0}^{2^n -1} (-1)^{f(x)}\ket{x}\frac{1}{\sqrt{2}}(\ket{0}-\ket{1}) \xrightarrow{H^{\otimes n}} \begin{cases} e^{i\pi
f(x)}\ket{0...0} \text{ if $f$ is constant} \\
\ket{y}, y \neq 0 \text{ if $f$ is balanced}
\end{cases}$$
Thus, if we measure the first register, we measure with certainty $\ket{0}$ if $f$ is constant and some other value if $f$ is balanced as desired. Once again, this scheme only queries the $f$-oracle once.

\subsection{Bernstein-Vazirani Algorithm}
We can use the Deutsch-Jozsa circuit to reveal hidden traits for another special class of boolean functions. Let $f\{0,1\}^n \rightarrow \{0,1\}$ be of the following form:
$$f(x) = a\cdot x \oplus b \mod 2$$ where $\oplus$ refers to bitwise binary addition and $\cdot$ refers to the dot product i.e $x \cdot y = \sum_{i=1}^n x_iy_i \mod 2$. $a \in \{0,1\}^n$ and $b \in \{0,1\}$ are hidden, and we wish to find these constants. In the classical world, we would have to query the oracle $\mathcal{O}(n)$ times to find both constants $a,b$. By using the Deutsch-Jozsa circuit, we can drop this to $\mathcal{O}(1)$ queries. We use the exact circuit above and calculate the effect of the circuit:

\begin{gather}
  \ket{0...0}\ket{1}\xrightarrow{H^{\otimes (n+1)}} \frac{1}{\sqrt{2^n}} \sum_{x=0}^{2^n-1}\ket{x}\frac{1}{\sqrt{2}}(\ket{0} - \ket{1}) \xrightarrow{U_f} \\
  \frac{1}{\sqrt{2^n}} \sum_{x=0}^{2^n-1}\ket{x}\frac{1}{\sqrt{2}}(\ket{f(x)} - \ket{1\oplus f(x)}) = \frac{1}{\sqrt{2^n}} \sum_{x=0}^{2^n-1}\ket{x}\frac{1}{\sqrt{2}}(\ket{a\cdot x \oplus b} - \ket{1 \oplus (a \cdot x \oplus b)}) = \\
  \frac{1}{\sqrt{2^n}} \sum_{x=0}^{2^n-1}(-1)^b \ket{x}\frac{1}{\sqrt{2}}(\ket{a\cdot x} - \ket{1 \oplus (a \cdot x)}) = \frac{(-1)^b}{\sqrt{2^n}} \sum_{x=0}^{2^n-1}(-1)^{a\cdot x}\ket{x}\frac{1}{\sqrt{2}}(\ket{0} - \ket{1})
\end{gather}

We prove a lemma showing destructive interference for certain terms: \newline

\begin{lemma} \label{dotprod}
  Let $z \in \{0,1\}^n$. Then
  $$\sum_{x=0}^{2^n-1} (-1)^{z \cdot x} = 0$$ iff $z \neq 0$
\end{lemma}
\begin{proof}
   We begin by expanding the form above:
   $$\sum_{x=0}^{2^n-1} (-1)^{z \cdot x} = \sum_{x=0}^{2^n -1}\prod_{j=1}^n (-1)^{z_jx_j} = \sum_{x_n = 0}^1...\sum_{x_2 = 0}^1\sum_{x_1 = 0}^1 \prod_{j=1}^n (-1)^{z_jx_j}   $$
   The last equality easily follows from the fact that we can rearrange the sum in the specific order imposed by the summation on the right.

   Now suppose there exists $1 \leq r \leq n$ such that $z_r = 1$. From the form above, we derive that:
   $$\sum_{x_n = 0}^1...\sum_{x_2 = 0}^1\sum_{x_1 = 0}^1 \prod_{j=1}^n (-1)^{z_jx_j} = \sum_{x_n = 0}^1...\sum_{x_{r+1} = 0}^1\sum_{x_{r-1} = 0}^1... \sum_{x_2 = 0}^1\sum_{x_1 = 0}^1 \prod_{j=1,j \neq r}^n (-1)^{z_jx_j} ((-1)^0 + (-1)^1) = 0 $$

   Note that from $\sum_{x=0}^{2^n-1} (-1)^{z \cdot x} \ket{x} = 2^n\delta_{z,0}$, we have the converse as well.
\end{proof}

We cannot use this lemma directly on the form representing the first $n$-qubit register above as the sum involves constituent states $\ket{x}$ tied to the $(-1)^{z\cdot x}$. To fix this, we first apply $H^{\otimes n}$ on the first register:

\begin{gather}
  \frac{(-1)^b}{\sqrt{2^n}} \sum_{x=0}^{2^n-1}(-1)^{a\cdot x}\ket{x} \xrightarrow{H^{\otimes n}} \frac{(-1)^b}{2^n}\sum_{x=0}^{2^n -1}  (-1)^{a\cdot x} \sum_{y=0}^{2^n -1} (-1)^{y \cdot x} \ket{y} = \\
  \frac{(-1)^b}{2^n}\sum_{y=0}^{2^n -1} \sum_{x=0}^{2^n -1} (-1)^{(a+y)\cdot x} \ket{y}
\end{gather}
%
We now apply Lemma \ref{dotprod} to the first register above to find that it equals:
$$(-1)^b \ket{a} $$ Thus, by measuring the first register, we get the value of $a$ off by some phase factor.

The main argument for the correctness of the algorithm stems from the destructive interference resulting from the operation of running an equal superposition of all $2^n$ possible bit-strings which can $x$ take, then ``merging" them together and letting destructive interference take its course.


\subsection{Grover's Algorithm}

Let $X$ be an unstructured database of size $N$. We can simply think of $X$ as an array with elements $\{1,...,N\}$ not necessarily in sorted order. Classically, we would have to query the database $\mathcal{O}(N)$ times. However, Grover's algorithm shows that it is possible to leverage quantum phenomena to bring this down to $\mathcal{O}(\sqrt{N})$ queries.

For the sake of a simpler analysis, let $f:\{0,1\}^n \rightarrow \{0,1\}^n$ be a boolean function representing an unstructured database of size $2^n$ with a single marked element $w \in \{0,n\}^n$.
$$ f(x) = \begin{cases}
            1 \text{ if } x = w \\
            0 \text{ otherwise}
          \end{cases} $$
The crux of the algorithm comes from iterating a specially-constructed unitary transformation $D$ composed with the oracle query gate $O$.

To begin, similar to the behavior found in the Deutsch-Jozsa, Bernstein-Vazirani algorithms, we exploit the phase-kickup technique by preparing the state $\ket{0^n} \otimes \ket{1}$, performing $O \circ(H^{\otimes n} \otimes H)$, then discarding the second register:

\begin{equation}\label{GrovOpAction}
\ket{0^n} \otimes \ket{1} \xrightarrow{\text{Query, Discard}} \frac{1}{\sqrt{2^n}} \sum_{x=0}^{2^n-1} (-1)^{f(x)} \ket{x}
\end{equation}
%
The second stage involves applying the below operator to the first $n$-qubit register.
%
$$ D = H^{\otimes n}(-I + 2\ket{0}\bra{0})H^{\otimes n} $$
%
Refer to $D$ as the \emph{Distillation Operator}. The composed operator $(D \otimes I) \circ U_f$ is denoted as the \emph{Grover Operator}. The middle term of $D$ is a conditional phase shift:

\[ T = (-I + 2\ket{0}\bra{0})\] with the following action:

\[ T\ket{0} = \ket{0}, \; T\ket{x} =-\ket{x} \text{for $  x \in \{1,\cdots, 2^n-1\}$}  \]

Let us compute the action of $D$ on $\ket{x} \in (\mathbb{C}^2)^{\otimes n}$ as follows: first compute the below expression:
%
\begin{gather}
  (-I + 2 \ket{0} \bra{0}) H^{\otimes n} \ket{x} = (-I + 2\ket{0}\bra{0})) \frac{1}{\sqrt{2^n}}\sum_{y=}^{2^n-1} (-1)^{y\cdot x}\ket{y} = \\
  \frac{1}{\sqrt{2^n}} \big[ -\sum_{y=0}^{2^n-1}(-1)^{y\cdot x}\ket{y} + 2\ket{0}\big]
\end{gather}
%
Now we apply the last $n$-qubit Hadamard transform to yield the following:
%
\begin{gather}
  \frac{1}{\sqrt{2^n}}H^{\otimes n} \big[-\sum_{y=0}^{2^n-1}(-1)^{y\cdot x}\ket{y} + 2\ket{0} \big] = \frac{1}{2^n}\big[-\sum_{z = 0}^{2^n-1} \sum_{y=0}^{2^n-1}(-1)^{y\cdot(x\oplus z)} \ket{y} + 2 \sum_{z = 0}^{2^n-1} \ket{z} \big] = \\
  \frac{1}{2^n}(-2^n\ket{x} + 2 \ket{S})
\end{gather}
%
where $\ket{S} = \sum_{z = 0}^{2^n-1} \ket{z}$.
%
The second-to-last equality follows from reindexing the sum with $y$-indices instead of $z$-indices. Note that the Hadamard transform is an inverse of itself, giving us $-\ket{x}$ as the first term.
%
Thus, we get that:
%
$$D\ket{x} = -\ket{x} + \frac{2}{2^n}\ket{S} $$
%
and we deduce the following $2^n\times 2^n$ matrix representation:
$$D = \left[ \begin{matrix}
        -1 + \frac{2}{2^n} & \frac{2}{2^n} & ... & \frac{2}{2^n} \\
        \frac{2}{2^n} & -1 + \frac{2}{2^n} & ... & \frac{2}{2^n} \\
        . & . &  ... & . \\
        . & . &  ... & . \\
        . & . &  ... & . \\
        \frac{2}{2^n} & \frac{2}{2^n} & ... & -1 + \frac{2}{2^n}
      \end{matrix} \right]
 $$
This operation is akin to the \emph{Inversion About the Mean} technique, which increases the amplitude of the solution $\ket{w} \otimes \ket{1}$ and deceases the amplitude of the others. To see this,  we find that the matrix $D$ acts on general quantum states as follows:

\begin{align*}
D(\sum_{z=0}^{2^n-1} \alpha_z\ket{z})
& = \sum_{z=0}^{2^n-1} \alpha_z (-\ket{z} + \frac{2}{2^n}\ket{S})
%
= -\sum_{z=0}^{2^n-1}\alpha_z \ket{z} + 2\sum_{k=0}^{2^n-1}\frac{\alpha_k}{2^n}\ket{S} \\
%
& = \sum_{z=0}^{2^n-1} \left( 2\sum_{k=0}^{2^n-1} \frac{\alpha_k}{2^n}   - \alpha_z \right)\ket{z}
%
 = \sum_{z=0}^{2^n-1} (2A -\alpha_z)\ket{z}
\end{align*}

The string of equalities show that the amplitudes of the state are reflected across the their mean.

\subsubsection{Geometric Interpretation}

The iteration of $DU_f$ has a geometric interpretation which we will now consider. Let $\ket{\psi}$ be the equally-weighted superposition and let $\ket{x_0}$ be the marked item:
$$ \ket{\psi} = \frac{1}{\sqrt{N}} \sum_{x=0}^{2^n-1} \ket{x} $$
%
We can express $\ket{\psi}$ as
$$ \ket{\psi} = \frac{1}{\sqrt{N}}\sum_{x \neq x_0} \ket{x} + \frac{1}{\sqrt{N}}\ket{x_0} = \sqrt{\frac{N-1}{N}}\ket{x_0^{\perp}} + \frac{1}{\sqrt{N}}\ket{x_0}$$
%
A crucial observation arises from the action of the Grover Operator $G$ on the subspace $T = \text{span}\{\ket{x_0}, \ket{x_0^{\perp}}\}$. Let us show that $G$ actually \emph{fixes} the subspace $T$. First,
the oracle operator $O$, has the following action on vector $ \alpha \ket{x_0^{\perp}} + \beta \ket{x_0}$,
\[ O(\alpha \ket{x_0^{\perp}} + \beta \ket{x_0}) = \alpha \ket{x_0^{\perp}} - \beta \ket{x_0}\]
This follows from the action discussed in \ref{GrovOpAction}. The application of the distillation operator to the vector $\alpha \ket{x_0^{\perp}} + \beta \ket{x_0}$ results in the following:

\[D(\alpha \ket{x_0^{\perp}} + \beta \ket{x_0}) = 2 \alpha \bra{\psi}\ket{x_0^{\perp}}\ket{\psi} + 2 \beta \bra{\psi}\ket{x_0}\ket{\psi}  - \alpha\ket{x_0^{\perp}} - \beta\ket{x_0} \]

Since $\ket{\psi} \in T$, the derivation shows that indeed $T$ is a $G$-invariant subspace. In fact, we can go even further and deduce that the distillation operator $D$ is actually a reflection, making $G$ a operator which rotates the starting state $\ket{\psi}$ counter-clockwise on the plane spanned by $\ket{x_0^{\perp}},\ket{x_0}$. It turns out that iterations of $G$ applied on $\ket{\psi}$ brings the output state closer to $\ket{x_0}$. However, we have to perform some analysis to ensure that our procedure doesn't ``overshoot" $\ket{x_0}$.

Let $\theta$ be the angle between $\ket{\psi}$ and $x_0^{\perp} = \frac{1}{\sqrt{N-1}}\sum_{x\neq x_0} \ket{x}$. This gives us the following representation:
$$ \ket{\psi} =  \cos{\theta}{x_0^{\perp}} + \sin{\theta}x_0 $$

Thus,
$$\sin{\theta} = \bra{x_0}\ket{\psi} = \frac{1}{\sqrt{N}}$$
implying that
$$ \theta \approx \frac{1}{\sqrt{N}} $$ for sufficiently large $N$.
We claim the following:
\begin{enumerate}
  \item $O$ corresponds to a reflection across hyperplane orthogonal to $\ket{x_0}$.
  \item $D$ corresponds to a reflection across hyperplane orthogonal to $\ket{\psi}$.
\end{enumerate}

These two operations rotate $\ket{\psi}$ by $2\theta$. To see this, imagine a generic unit vector $\ket{\Psi}$ some angle $\theta_0$ from $x_0^{\perp}$ on the two-dimensional plane spanned by $\ket{x_0}$ and $\ket{x_0^{\perp}}$. Assume that we orient our plane such that we visualize $\ket{x_0^{\perp}}$ as our horizontal axis and $\ket{x_0}$ as our vertical axis. Reflecting across the axis $\ket{x_0^{\perp}}$ rotates our $\ket{\Psi}$ $2\theta_0$ clockwise, giving position $-\theta_0$. Then reflecting it across axis $\ket{S}$ rotates it by $2(\theta_0 + \theta)$, giving the angle $2(\theta_0 + \theta) - \theta_0 = \theta_0 + 2\theta$.

Since $\theta \approx \frac{1}{\sqrt{N}}$ is close to zero for sufficiently large $N$, it suffices to rotate our equally-weighted superposition $\ket{S}$ around $\frac{\pi}{2}$ to become close to $\ket{x_0}$ as $\ket{x_0^{\perp}},\ket{x_0}$ are orthogonal to each other. We thus solve the following:
$$2k\theta \approx \frac{\pi}{2} \implies k = \frac{\sqrt{N}\pi}{4} $$

This gives us that we need about $\mathcal{O}(\sqrt{N})$ iterations to give us a state when measured gives us the desired index with high probability. As each iteration queries the oracle a constant number of times, it follows that we need $\mathcal{O}(\sqrt{N})$ queries.

\subsection{Quantum Fourier Transform}

We now investigate the quantum analogue of the discrete Fourier transform as follows:
%
For an $n$-qubit system, we define the \textit{Quantum Fourier Transform}:
%
$$ F = \frac{1}{\sqrt{2^n}} \sum_{x,y = 0}^{2^n-1} \zeta^{xy} \ket{y}\bra{x} $$
 where $\zeta = e^{2\pi i /2^n} $.
%
 This operation is unitary as we can directly check:
 $$ F^{\dagger}F = \frac{1}{2^n} \sum_{y = 0}^{2^n-1} \sum_{x = 0}^{2^n-1} \zeta^{xy - xy} \ket{x}\bra{x} = \sum_{x = 0}^{2^n -1} \ket{x}\bra{x} = I $$
%
 We can explictly express the action of $F$ on a basis vector $x \in (\mathbb{C}^2)^{\otimes n}$
\begin{gather*}
  \begin{split}
    F\ket{x} & = \frac{1}{\sqrt{2^n}} \sum_{y=0}^{2^n-1} \zeta^{xy} \ket{y} \\
     & = \frac{1}{\sqrt{2^n}} \sum_{y_0=0}^{1}\sum_{y_2=0}^{1}...\sum_{y_{n-1}=0}^{1} \zeta^{x(\sum_{j=1}^n y_j2^j)} \ket{y_{n-1}} ...\ket{y_0} \\
     & = \frac{1}{\sqrt{2^n}} \sum_{y_0=0}^{1}\sum_{y_1=0}^{1}...\sum_{y_{n-1}=0}^{1} \prod_{j=0}^{n-1} \zeta^{xy_j2^j} \ket{y_{n-1}} \ldots \ket{y_0}  \\
    & = \frac{1}{\sqrt{2^n}} \bigotimes_{i=0}^{n-1}\sum_{y_i=0}^1 \zeta^{xy_i2^i} \ket{y_i} \\
    & = \frac{1}{\sqrt{2^n}} \bigotimes_{i=0}^{n-1}\ket{0} + \zeta^{x2^i}\ket{1} \\
    & = \frac{1}{\sqrt{2^n}} \bigotimes_{i=1}^{n-1}\ket{0} + \zeta^{\sum_{k=0}^{n-1} x_k2^{k+i}}\ket{1} \\
    & = \frac{1}{\sqrt{2^n}} \bigotimes_{i=0}^{n-1}\ket{0} + e^{2\pi i(x_{n-1}2^{i-1} +x_{n-2}2^{i-2} + \ldots + x_{0} 2^{i-n}))}\ket{1} \\
    & = \frac{1}{\sqrt{2^n}} \bigotimes_{i=0}^{n-1}\ket{0} + e^{2\pi i(x_{n- i - 1}2^{-1} +x_{n- i - 2}2^{-2} + \ldots + x_{0} 2^{i-n})}\ket{1} \\
    & = \frac{1}{\sqrt{2^n}} \bigotimes_{i=0}^{n-1}\ket{0} + e^{2\pi i [ 0.x_{n-i-1}\ldots x_1x_0 ]} \ket{1}
  \end{split}
\end{gather*}
%
The last equality follows from the fact that any bit shift past $n-1-k$ times will result in an even power of $e^{2\pi i}$ which will always be equal to one.
where
$$P_i = \left[\begin{matrix} 1 & 0 \\ 0 & \zeta^{2^i} \end{matrix}\right]$$ is the phase shift gate rotating $\zeta^{2^i}$ Note that
%
$$ e^{2\pi i(x_{n-1-i}{2^{-1})}} = e^{\pi i x_{n-1 - i}} =
\begin{cases}
0 \text{ if } x_{n - 1 - i} = 0 \\
-1 \text{ if } x_{n - 1 - i} = 1
\end{cases}
$$
%
Thus, $\ket{0} + e^{2\pi i(x_{n-1-i}{2^{-1})}}\ket{1}$ can be realized through a Hadamard transform on each $\ket{x_{n_ 1- i}}, \newline \; 0 \leq i \leq n-1$.
%
Now from this expression, we can construct the quantum circuit implementing the Quantum Fourier Transform for a basis vector $x$ in Figure \ref{fig:QFT}.
\begin{figure}[h]
  \Qcircuit @C=0.6em @R=1.2em{
  \lstick{\ket{x_0}} & \qw & \qw & \qw & \qw & \ldots & & & \qw & \ctrl{8} & \qw  & \qw & \qw & \qw & \cdots & & & \ctrl{7} & & \qw & \cdots & & \ctrl{1} & \qw & \qw & \push{\ket{z_{n-1}}} \\
  \lstick{\ket{x_1}} & \qw & \qw & \qw & \qw & \ldots & & &  \ctrl{7} & \qw & \qw & \qw & \qw & \qw & \cdots & & \ctrl{6} & \qw & \qw & \cdots & & & \gate{H} & \gate{P_2} & & \qw & \ket{z_{n-2}} \\
  . \\
  . \\
  . \\
  \lstick{\ket{x_{n-4}}} & \qw & \qw & \qw & \ctrl{3} & \cdots & & & \qw & \qw & \qw & \qw & \ctrl{2} & \cdots & & &\qw & \qw & \qw  & \cdots & & &\qw & \qw & \qw & \qw & \qw & \ket{z_{3}} \\
  \lstick{\ket{x_{n-3}}} & \qw & \qw & \ctrl{2} &\qw  &\ldots & &  & \qw & \qw & \qw & \ctrl{1} & \qw & \cdots & & &\qw & \qw & \qw  & \cdots & & & & \qw & \qw & \qw & \qw & \ket{z_2} \\
  \lstick{\ket{x_{n-2}}} & \qw & \ctrl{1} & \qw & \qw &\cdots & & & \qw  & \qw & \gate{H} & \gate{P_2} & \gate{P_3} & \cdots & & &\gate{P_{n-2}} & \gate{P_{n-1}} & \qw & \cdots & & & \qw & \qw & \qw & \qw & \ket{z_1} \\
  \lstick{\ket{x_{n-1}}} & \gate{H} & \gate{P_2} & \gate{P_3} & \gate{P_4} &\cdots & & & \gate{P_{n-1}} & \gate{P_n} & \qw & \qw & \qw & \cdots & & & \qw & \qw & \qw & \cdots & & & \qw & \qw & \qw & \qw & \ket{z_0}
  }
  \caption{Quantum circuit realizing the QFT}
  \label{fig:QFT}
\end{figure}

\subsection{Phase Estimation}

The setup of the phase estimation problem is the following: one is given an unitary operator $U$ and a state $\ket{\phi}$ promised to be an eigenvector of $U$ such that
\begin{equation}
  U\ket{\phi} = e^{i\phi}\ket{\phi}
\end{equation}
%
We wish to determine an $n$-bit estimate of $\phi$. The quantum circuit which achieves this is as below:
%
\begin{figure}[h]
  \Qcircuit @C=0.6em @R=1.2em {
  \lstick{\ket{0}} & \gate{H} & \qw & \qw & \qw & & \ldots & & \qw & \\
  \lstick{\ket{0}} & \gate{H} & \qw & \qw & \qw & & \ldots & & \qw & \\
  \lstick{.} \\
  \lstick{.}  \\
  \lstick{.}  \\
  \lstick{\ket{0}} & \gate{H} & \qw & \ctrl{2} & \qw & &\ldots & & \qw &\\
  \lstick{\ket{0}} & \gate{H} & \ctrl{1} & \qw & \qw  & &\ldots & & \qw &\\
  \lstick{\ket{\phi}} & \qw & \gate{U} & \gate{U^2} & \qw & &\ldots & &
  }
  \caption{Phase Estimation Circuit}
\end{figure}
%
We first prepare the initial quantum state as the uniform superposition:
$$ \frac{1}{\sqrt{2^n}}\sum_{x-0}^{2^n -1} \ket{x} \otimes \ket{\phi} $$
%
applying the operator
\begin{equation}
  \sum_{x=0}^{2^n -1} \ket{x}\bra{x} \otimes U^x
\end{equation}
to the initial state gives way to:
%
$$ \sum_{x = 0}^{2^n - 1} e^{i\phi x}\ket{x} \otimes \ket{\phi} $$
Equivalently this state can be written as:
\begin{equation}
    \sum_{x = 0}^{2^n-1} e^{\frac{\phi}{2 \pi} [2 \pi i x ] } \ket{x} \otimes \ket{\phi}
\end{equation}
Suppose that $\phi/2 \pi$ terminates after at most $n$-bits or if $\frac{\phi}{2\pi} = y/2^n$ for some $y \in \mathbb{Z}_{2^n}$.
%
Then $\phi = \frac{2\pi y}{2^n}$. Substituting this in the expression above and applying the inverse QFT to the first $n$ registers will result in the state $\ket{y} \otimes \phi$. Therefore, measuring on the first register will give us the exact value of $y$.

(Todo: think about the general case)

\newcommand{\R}{{\bf R}}
\newcommand{\A}{{\bf A}}

\subsection{Watrous' Algorithm for Solvable Groups}
Watrous showed that recovering the generators and their orders for the case of solvable groups can be computed efficiently using a quantum computer \cite{watrous2001quantum}. Recall that the \emph{commutator subgroup} of $G$ is defined as
\[ G' = [G,G] = \{ghg^{-1}h^{-1} \mid g,h \in G \} \]
%
From the observation that $[G,G]$ is actually normal in $G$, construct the tower of subgroups
\[ G \trianglerighteq G' \trianglerighteq  (G')' \trianglerighteq  \cdots \]
named the \emph{derived series} of $G$. This tower is not guaranteed to be finite for general groups. However, groups whose derived series collapses at $\{e\}$ are deemed \emph{solvable}
%
\[ \{e\} = N_0 \trianglelefteq N_1 \trianglelefteq N_2 \cdots \trianglelefteq N_m = G \]
%
An equivalent characterization of solvable group $G$ is if there exist a set of generators $g_1,\cdots, g_m \in G$ and finitely-generated subgroups $H_j = \langle g_1,\cdots, g_j \rangle$ such that $G$ can be expressed as a composition series $\{e\} = H_0 \trianglelefteq H_1 \trianglelefteq \cdots \trianglelefteq H_m = G$ with the $H_j / H_{j-1}$ being cyclic $p$-groups for all such $j$. In other words, $H_j = \bigsqcup_{i = 0}^{\alpha_j} g_j^{i} H_j$ where some power of a prime $p$. See \cite{aluffi2009algebra} for more information about composition series and their refinements.

Babai and Beals presented a classical Monte-Carlo algorithm to determine the generators in polynomial time \cite{babai1995fast} for finite groups. With this baseline, it suffices to determine the orders of the generators $\alpha_j$. To accomplish this, we will need access to uniform superpositions of the $H_j$ to perform a close variant of Shor's algorithm. Watrous' algorithm is divided into two phases for step $j$. Assume that we have access to the uniform superposition
%
\[\ket{H_{j-1}} = \frac{1}{\sqrt{|H_{j-1}}|} \sum_{h \in H_{j-1}} \ket{h} \]
%
For ease of notation, we will denote $H_{j-1}$ as $H$, the generator $g_{j}$ as $g$, and the order of $g$, $\alpha_j$, as $\alpha$. Furthermore, assume that we have determined $m$ generators for our finite group $G$. The algorithm utilizes a large number of registers which are defined $\R$ or $\A$ registers whose definitions will be fleshed out in the coming sections.

\subsubsection{Phase One of Step $j$}
For this phase, we assume we have access to a uniform superposition $\ket{H}$ initialized in a quantum register $\R$. We can think of $\R$ just as some finite dimensional complex Hilbert space with an orthogonal basis indexed by the group elements of $G$. Furthermore, our $\A$ register will have an orthogonal basis index by $\mathbb{Z}_M$ for some sufficiently large $M$ to be determined later on. The $\A$ is initialized to $\ket{0}$.
%
Suppose we also have access to the unitary operators $M: \R \otimes \R \rightarrow \R \otimes \R$ and $V_g: \A \otimes
\R \rightarrow \R$ with the following actions on basis vectors:
%
\begin{align*}
  & M\ket{g} \otimes \ket{h} = \ket{g} \otimes \ket{gh} \quad g,h \in G \\
  & V_g \ket{a} \otimes \ket{h} = \ket{g^ah} \quad b \in \mathbb{Z}_M, h \in G
\end{align*}
%
We now begin with the entangled state $\ket{0} \otimes \ket{H} \in \A \otimes \R$. The operations below yield the following:
%
\begin{gather*}
\ket{0} \otimes \ket{H} \xrightarrow{QFT_{M} \otimes I_\R} \frac{1}{\sqrt{M}} \sum_{a \in \mathbb{Z}_M} \ket{a} \otimes \ket{H} \xrightarrow{V_g} \frac{1}{\sqrt{M}} \sum_{a \in \mathbb{Z}_M} \ket{a} \otimes \ket{g^aH} \\
\xrightarrow{QFT^{-1}_M \otimes I_\R} \frac{1}{M}  \sum_{a \in \mathbb{Z}_M} \sum_{b \in \mathbb{Z}_M} e^{-2\pi ab/M}\ket{b} \otimes \ket{g^aH}
\end{gather*}

The observation of the $\A$ register will lead to an analysis similar to that of Shor's algorithm. From \cite{watrous2001quantum}, it turns out that setting $M = 2^{2n + \mathcal{O}(1/\epsilon)}$ suffices for our $\A$ register dimension.

(Fill this part in).

\subsubsection{Phase Two of Step $j$}
We now need a method to generate uniform superpositions of $H_j$ from those of $H_{j-1}$. We can view these uniform superpositions as resources for sampling required in Phase One through Shor's Algorithm. Assuming we have access to $\ell$ copies of $\ket{H_{j-1}}$, we can generate $\ell - 1$ copies of $H_{j}$ for the next step using a clever trick outlined below. First, we note that the orders $\alpha_i$ have been

Let $\R_1, \R_2, \cdots \R_\ell$ denote the copies of $\ket{H_{j-1}}$ and $\A_1, \A_2, \cdots \A_\ell$ the corresponding $\A$ registers initialized to $\ket{0}$. Once again, we denote the uniform superposition $\ket{H_{j-1}}$ as $\ket{H}$ and label the variables $\alpha_i, g_i$ as $\alpha, g$ respectively.

We will be considering the entangled pairs $\A_i \otimes \R_i$ for $1 \leq i \leq \ell$ and perform the following series of unitary operations:

\begin{align*}
  \ket{0} \otimes \ket{H}
  & \xrightarrow{QFT^{-1} \otimes I_{\bf R_i}} \frac{1}{\sqrt{\alpha}} e^{2\pi a_i/\alpha} \ket{a_i} \otimes \ket{H} \xrightarrow{V_g} \frac{1}{\sqrt{\alpha}}\sum_{a_i \in \mathbb{Z_\alpha}} \ket{a_i} \otimes \ket{g^{a_i}H} \\
  & \xrightarrow{QFT \otimes I_{\bf R_i}} \frac{1}{\alpha}\sum_{a_i \in \mathbb{Z}_\alpha}\sum_{b_i \in \mathbb{Z}_\alpha} e^{2\pi i a_ib_i } \ket{b_i} \otimes \ket{g^{a_i}H}
\end{align*}
%
Now by measuring the ${\bf A_i}$ register and discarding it, we get the following state with uniform probability and the measurement $b_i$:
%
\[\ket{\psi_i} = \frac{1}{\sqrt{r}} \sum_{a_i \in \mathbb{Z}_\alpha} e^{2\pi i a_i b_i / \alpha} \ket{g^{a_i}H} \]
%
Repeat this for all ${\bf A_i} \otimes {\bf R_i}$ yield the measurements $b_1, \cdots, b_\ell$ and states $\ket{\psi_1}, \cdots, \ket{\psi_\ell}$ for all $1 \leq i \leq \ell$.
%
So all that really remains is to get rid of the phase in front of the state somehow. It turns out that you can use a state $\ket{\psi_k}$ where $b_k$ relatively prime to $b_\ell$ to remove phase for all $\ket{\psi_i}$ for $i \neq k$. To see this, assume that we have found such a $b_k$ in order to consider the state in ${\bf R_i} \otimes {\bf R_k}, \ket{\psi_i} \otimes \ket{\psi_k}$, and the unitary operator $M_{(g^{a_i}h)^c}$ with the action:
%
\[ M_{(g^{a_i})^ch} \ket{z} \rightarrow  \ket{(g^{a_i})^chz}, \quad z \in G\]
%
We first show that $\ket{\psi_k}$ is an eigenvector of the $M_{g^{a_i}h}$ operator:
\begin{align*}
M_{g^{a_i}h} \ket{\psi_k}
& = \frac{1}{\sqrt{r}} \sum_{a_k \in \mathbb{Z}_\alpha} e^{2\pi i a_kb_k / \alpha} M_{g^{a_i}h} \ket{g^{a_k}H} =  \frac{1}{\sqrt{r}|H|} \sum_{a_k \in \mathbb{Z}_\alpha} \sum_{h' \in H} e^{2\pi i a_kb_k / \alpha} M_{g^{a_i}h} \ket{g^{a_k}h'} \\
& = \frac{1}{\sqrt{r}|H|} \sum_{a_k \in \mathbb{Z}_\alpha} \sum_{h' \in H} e^{2\pi i a_kb_k / \alpha} \ket{g^{a_i+a_k}h''h'} = \frac{1}{\sqrt{r}|H|} \sum_{a_k \in \mathbb{Z}_\alpha} \sum_{h' \in H} e^{2\pi i (a_k-a_i)b_k / \alpha} \ket{g^{a_k}h'} \\
& =  e^{-2\pi i a_ib_k}  \frac{1}{\sqrt{r}} \sum_{a_k \in \mathbb{Z}_\alpha} e^{2\pi i a_kb_k / \alpha} \ket{g^{a_k}H} \\
& = e^{-2\pi i a_ib_k} \ket{\psi_k}
\end{align*}
%
The third equality follows from the normality of $H$ as there must exist some $h'' \in H$ such that $hg^{a_i} = g^{a_i}h''$.
%
Using this fact, we can easily show that simply composing the multiplication operator on the state $\ket{\psi_i} \otimes \ket{\psi_k}$ $c$ times where $c = b_ib_k^{-1} \mod \alpha$. Such a $c$ must exist since $b_k$ is relatively prime to $r$. Now observe the following unitary transformations:

\begin{align*}
\ket{\psi_i} \otimes \ket{\psi_k}
& = \frac{1}{\sqrt{r}}\sum_{a_i \in \mathbb{Z}_\alpha} \sum_{h \in H} e^{2 \pi i a_i b_i / \alpha} \ket{g^{a_i}h} \otimes \ket{\psi_k} \xrightarrow{M^c}  \frac{1}{\sqrt{r}}\sum_{a_i \in \mathbb{Z}_\alpha} \sum_{h \in H} e^{2 \pi i a_i b_i / \alpha} \ket{g^{a_i}h} \otimes M_{(g^{a_i}h)^c}\ket{\psi_k} \\
& = \frac{1}{\sqrt{r}}\sum_{a_i \in \mathbb{Z}_\alpha} \sum_{h \in H} e^{2 \pi i a_i b_i / \alpha} \ket{g^{a_i}h} \otimes e^{-2\pi i ca_ib_k}\ket{\psi_k} = \frac{1}{\sqrt{r}}\sum_{a_i \in \mathbb{Z}_\alpha} \sum_{h \in H} \cancel{e^{2 \pi i a_i b_i / \alpha}} \ket{g^{a_i}h} \otimes \cancel{e^{-2\pi ia_ib_i}}\ket{\psi_k} \\
& = \ket{\langle g \rangle H} \otimes \ket{\psi_{k}}
\end{align*}
%
Thus, with $\ket{\psi_k}$, we can turn $\ell$ $\R$ registers initialized to $\ket{H}$ to $\ell -1$ registers set to $\ket{\langle g \rangle H} = \ket{H_j}$. The algorithm can now proceed to the next step.
%
We now outline the full algorithm for the finite cyclic tower $\{e\} = H_0 \trianglelefteq H_1 \trianglelefteq H_2 \cdots \trianglelefteq H_m = G$: \newline

\hrule
\begin{enumerate}
  \item Prepare $(k+1)(m+1)$ $\R$ registers in state $\ket{H_0}$ where $k$ is the number of trials required to output the correct order through Shor's Algorithm at every step with high probability.
  \item Perform phase one of the algorithm to find the orders $\alpha_j$ at step $j$ using $k$ copies of $\ket{H_{j-1}}$.
  \item Use one of the copies of $\ket{H_{j-1}}$ to turn the remaining copes of $\ket{H_{j-1}}$ into $\ket{H_{j}}$ through phase two.
  \item Output the $\alpha_1, \cdots, \alpha_m$
\end{enumerate}
\hrule

Note that here we tacitly assume that we have a commensurate number of $\A$ registers all initialzed to $\ket{0}$ with the proper dimensions required in phase one and two.

\begin{tcolorbox}
  \begin{itemize}
    \item I wonder if we can use this technique over finite fields $\mathbb{F}_p$ for some prime power $p$ to compute some interesting objects studied in Galois Theory
    \item Is there a way to extend these algorithms to find generators for intersections of subgroups. Could be interested to find an algorithm to calculate refinements through Schrier refinement theorem, Zassenhaus' Butterfly lemma.
  \end{itemize}

\end{tcolorbox}




%There are $m$ $R$ registers $\R_1,\R_2, \cdots, \R_m$ and $\A$ registers $\A_1, \A_2, \cdots, \A_m$. The $j^{th}$ step utilizes the $\R_j, \A_j$ registers where $\R_j$ has orthogonal basis vectors indexed by group elements
